---
title: "NUMS Examples"
author: "Jaylin Lowe and Jack Moran"
date: "10/11/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

We will use the classroom dataset (West et. al, 2014). 
```{r}
data("classroom", package = "WWGbook")
```

We will fit a simple three-level hierarchical linear model using the `lme4` package: 
```{r}
class.lmer <- lme4::lmer(mathgain ~ mathkind + sex + minority + ses + housepov + (1|schoolid/classid), data = classroom)
summary(class.lmer)
```

##Residual Analysis

We use the new function, `hlm_resid`, to calculate level-1 residuals for the classroom model.
```{r}
(resid_1 <- hlm_resid(class.lmer, 
                      level = 1, 
                      standardize = FALSE))
```

Note that warning that sample size for LS residuals is insufficient, so we have to use EB level-1 residuals. We plot the residuals versus explanatory variables or fitted value.
```{r}
ggplot(data = resid_1, aes(x = mathkind, y = .resid)) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```

We continue with level-2 residuals:
```{r}
(resid_2 <- hlm_resid(class.lmer, 
                      level = "classid:schoolid", 
                      standardize = FALSE, 
                      include.ls = FALSE))

ggplot(data = resid_2, aes(x = classid, y = .ranef.intercept)) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```

And level-3 residuals:
```{r}
(resid_3 <- hlm_resid(class.lmer, 
                      level = "schoolid", 
                      standardize = FALSE,
                      include.ls = FALSE))
```

##Influence Analysis 

Using the new function `hlm_influence`, we can calculate influence diagnostics for every student in the dataset: 
```{r}
(infl <- hlm_influence(class.lmer, level = 1, leverage = "overall"))
```

```{r}
dotplot_diag(infl$cooksd, name = "cooks.distance", cutoff = "internal")
```

Since this is a hierarchical model, we can also do this at the second level (classes), or the third (schools):
```{r}
#second level 
(infl2 <- hlm_influence(class.lmer, level = "classid:schoolid", cutoff = "internal"))
dotplot_diag(infl2$cooksd, name = "cooks.distance", cutoff = "internal")

#third level 
(infl3 <- hlm_influence(class.lmer, level = "schoolid", cutoff = "internal"))
dotplot_diag(infl3$cooksd, name = "cooks.distance", cutoff = "internal")
```

We can also use the function `hlm_augment` to return both influence diagnostics and residuals. 
```{r}
aug <- hlm_augment(class.lmer, level = 1)
print(aug, width = Inf)
```

